---
title: "Prediction Algorithm using Network Statistics"
author: "David Jia, David Keller, Alvaro Gonzalez"
date: "May 17, 2018"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

The igraph library is necessary.

```{r collapse=TRUE, warning=FALSE}
library(igraph)
library(pROC)
library(rpart)
library(rpart.plot)
```

The data set used was a Speed Dating one.

```{r}
dat = read.csv("https://raw.githubusercontent.com/David-Keller/DAiG/master/ADiG/SpeedDatingData.csv")
```

```{r, echo=FALSE}
# add zero imputation
impute_zero = function(x) {
  stopifnot(length(x) > 0 && sum(!is.na(x)) > 0)
  x[is.na(x)] = 0
  return(x)
}

# add negative one imputation
impute_negative_one = function(x) {
  stopifnot(length(x) > 0 && sum(!is.na(x)) > 0)
  x[is.na(x)] = -1
  return(x)
}

# add mean imputation
impute_mean = function(x) {
  stopifnot(length(x) > 0 && sum(!is.na(x)) > 0)
  x[is.na(x)] = median(x, na.rm=TRUE)
  return(x)
}

split_data = function(dat, frac=c(0.75, 0.25)) {
  # at least one set must be specified
  k = length(frac)
  stopifnot(k > 0)
  
  n = nrow(dat)
  frac = frac/(sum(frac))
  starts = c(1, round(cumsum(frac) * n)[-k])
  ends = c(starts[-1]-1,n)
  samp = sample(1:n)
  data_sets = list()
  for (i in 1:k) {
    data_sets[[i]] = dat[samp[starts[i]:ends[i]],]
  }
  return(data_sets)
} 

# Find most common value, return first value if tie
Mode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}

layout.by.attr <- function(graph, wc, cluster.strength=1,layout=layout.auto) {  
  g <- graph.edgelist(get.edgelist(graph)) # create a lightweight copy of graph w/o the attributes.
  E(g)$weight <- 1
  
  attr <- cbind(id=1:vcount(g), val=wc)
  g <- g + vertices(unique(attr[,2])) + igraph::edges(unlist(t(attr)), weight=cluster.strength)
  
  l <- layout(g, weights=E(g)$weight)[1:vcount(graph),]
  return(l)
}
```

### Creating Second Data Set

Filling in some of the missing data from the original data set and create a second data set that has the attribute of each individual node. The orgiinal data set had attributes from the pairings of two nodes.

```{r}
dat$age = impute_mean(dat$age)
dat$age_o = impute_mean(dat$age_o)
dat$race = impute_zero(dat$race)
dat$imprace = impute_mean(dat$imprace)
dat$sports = impute_mean(dat$sports)
dat$tvsports = impute_mean(dat$tvsports)
dat$exercise = impute_mean(dat$exercise)
dat$dining = impute_mean(dat$dining)
dat$museums = impute_mean(dat$museums)
dat$hiking = impute_mean(dat$hiking)
dat$gaming = impute_mean(dat$gaming)
dat$clubbing = impute_mean(dat$clubbing)
dat$reading = impute_mean(dat$reading)
dat$tv = impute_mean(dat$tv)
dat$theater = impute_mean(dat$theater)
dat$movies = impute_mean(dat$movies)
dat$concerts = impute_mean(dat$concerts)
dat$music = impute_mean(dat$music)
dat$shopping = impute_mean(dat$shopping)
dat$yoga = impute_mean(dat$yoga)


dat2 = data.frame(iid=unique(dat$iid))
dat2$age = aggregate(age ~ iid, data = dat, mean)$age
dat2$gender = aggregate(gender ~ iid, data = dat, mean)$gender
dat2$race = aggregate(race ~ iid, data = dat, mean)$race
dat2$hasMatch = aggregate(match ~ iid, data = dat, max)$match
dat2$metWithSameRace = aggregate(samerace ~ iid, data = dat, max)$samerace
dat2$imprace = aggregate(imprace ~ iid, data = dat, mean)$imprace
dat2$sports = aggregate(sports ~ iid, data = dat, mean)$sports
dat2$tvsports = aggregate(tvsports ~ iid, data = dat, mean)$tvsports
dat2$exercise = aggregate(exercise ~ iid, data = dat, mean)$exercise
dat2$dining = aggregate(dining ~ iid, data = dat, mean)$dining
dat2$museums = aggregate(museums ~ iid, data = dat, mean)$museums
dat2$hiking = aggregate(hiking ~ iid, data = dat, mean)$hiking
dat2$gaming = aggregate(gaming ~ iid, data = dat, mean)$gaming
dat2$clubbing = aggregate(clubbing ~ iid, data = dat, mean)$clubbing
dat2$reading = aggregate(reading ~ iid, data = dat, mean)$reading
dat2$tv = aggregate(tv ~ iid, data = dat, mean)$tv
dat2$theater = aggregate(theater ~ iid, data = dat, mean)$theater
dat2$movies = aggregate(movies ~ iid, data = dat, mean)$movies
dat2$concerts = aggregate(concerts ~ iid, data = dat, mean)$concerts
dat2$music = aggregate(music ~ iid, data = dat, mean)$music
dat2$shopping = aggregate(shopping ~ iid, data = dat, mean)$shopping
dat2$yoga = aggregate(yoga ~ iid, data = dat, mean)$yoga
```

```{r, echo=FALSE}
iidOfSameRace = dat[dat$samerace == 1,]$iid
pidOfSameRace = dat[dat$samerace == 1,]$pid

edges = c()

i = 0
while (i < length(iidOfSameRace)) {
  iid = iidOfSameRace[i+1]
  pid = pidOfSameRace[i+1]
  if (iid > 117) {
    iid = iid - 1
  }
  if (is.na(pid)) {
    pid = iid
  }
  if (pid > 117 & !is.na(pid)) {
    pid = pid - 1
  }
  edges = c(edges, iid)
  edges = c(edges, pid)
  i = i + 1
}
```

### Graph

The graph created here shows the connections between nodes where they met with someone of the same race.

```{r, echo=FALSE}
g = make_empty_graph(n = length(unique(dat$iid)), directed=FALSE) %>% add_edges(edges)
g = simplify(g, remove.multiple=TRUE)
V(g)$label = unique(dat$iid)
plot(g, vertex.size=2, edge.width=1, edge.color="red", vertex.label.cex=0.5, vertex.label.dist=0.4, vertex.label.degree=-pi/2, layout=layout.by.attr(g, wc=1))
```

The number of nodes in each component was recorded and assigned to the respective nodes in the data set.

```{r}
compDF = data.frame(components(g)$membership)
compTableDF = data.frame(table(components(g)$membership))
nameMatchVar = setNames(as.character(compTableDF$Freq), compTableDF$Var1)
totalCompDF = data.frame(lapply(compDF, function(i) nameMatchVar[i]))
componentTotalNodes = totalCompDF$components.g..membership
dat2$componentNumber = compDF$components.g..membership
averageAge = aggregate(age ~ componentNumber, data=dat2, mean)
nameMatchVar2 = setNames(as.character(averageAge$age), averageAge$componentNumber)
averageAgeDF = data.frame(lapply(compDF, function(i) nameMatchVar2[i]))
averageAgeOfComponent = averageAgeDF$components.g..membership
percentOfMales = aggregate(gender ~ componentNumber, data=dat2, mean)
nameMatchVar3 = setNames(as.character(percentOfMales$gender), percentOfMales$componentNumber)
percentOfMalesDF = data.frame(lapply(compDF, function(i) nameMatchVar3[i]))
percentOfMalesInComponent = percentOfMalesDF$components.g..membership
averageImprace = aggregate(imprace ~ componentNumber, data=dat2, mean)
nameMatchVar4 = setNames(as.character(averageImprace$imprace), averageImprace$componentNumber)
averageImpraceDF = data.frame(lapply(compDF, function(i) nameMatchVar4[i]))
averageImpraceOfComponent = averageImpraceDF$components.g..membership
averageSports = aggregate(sports ~ componentNumber, data=dat2, mean)
nameMatchVar5 = setNames(as.character(averageSports$sports), averageSports$componentNumber)
averageSportsDF = data.frame(lapply(compDF, function(i) nameMatchVar5[i]))
averageSportsOfComponent = averageSportsDF$components.g..membership
averageTvsports = aggregate(tvsports ~ componentNumber, data=dat2, mean)
nameMatchVar6 = setNames(as.character(averageTvsports$tvsports), averageTvsports$componentNumber)
averageTvsportsDF = data.frame(lapply(compDF, function(i) nameMatchVar6[i]))
averageTvsportsOfComponent = averageTvsportsDF$components.g..membership
averageExercise = aggregate(exercise ~ componentNumber, data=dat2, mean)
nameMatchVar7 = setNames(as.character(averageExercise$exercise), averageExercise$componentNumber)
averageExerciseDF = data.frame(lapply(compDF, function(i) nameMatchVar7[i]))
averageExerciseOfComponent = averageExerciseDF$components.g..membership
averageDining = aggregate(dining ~ componentNumber, data=dat2, mean)
nameMatchVar8 = setNames(as.character(averageDining$dining), averageDining$componentNumber)
averageDiningDF = data.frame(lapply(compDF, function(i) nameMatchVar8[i]))
averageDiningOfComponent = averageDiningDF$components.g..membership
averageMuseums = aggregate(museums ~ componentNumber, data=dat2, mean)
nameMatchVar9 = setNames(as.character(averageMuseums$museums), averageMuseums$componentNumber)
averageMuseumsDF = data.frame(lapply(compDF, function(i) nameMatchVar9[i]))
averageMuseumsOfComponent = averageMuseumsDF$components.g..membership
averageHiking = aggregate(hiking ~ componentNumber, data=dat2, mean)
nameMatchVar10 = setNames(as.character(averageHiking$hiking), averageHiking$componentNumber)
averageHikingDF = data.frame(lapply(compDF, function(i) nameMatchVar10[i]))
averageHikingOfComponent = averageHikingDF$components.g..membership
averageGaming = aggregate(gaming ~ componentNumber, data=dat2, mean)
nameMatchVar11 = setNames(as.character(averageGaming$gaming), averageGaming$componentNumber)
averageGamingDF = data.frame(lapply(compDF, function(i) nameMatchVar11[i]))
averageGamingOfComponent = averageGamingDF$components.g..membership
averageClubbing = aggregate(clubbing ~ componentNumber, data=dat2, mean)
nameMatchVar12 = setNames(as.character(averageClubbing$clubbing), averageClubbing$componentNumber)
averageClubbingDF = data.frame(lapply(compDF, function(i) nameMatchVar12[i]))
averageClubbingOfComponent = averageClubbingDF$components.g..membership
averageReading = aggregate(reading ~ componentNumber, data=dat2, mean)
nameMatchVar13 = setNames(as.character(averageReading$reading), averageReading$componentNumber)
averageReadingDF = data.frame(lapply(compDF, function(i) nameMatchVar13[i]))
averageReadingOfComponent = averageReadingDF$components.g..membership
averageTv = aggregate(tv ~ componentNumber, data=dat2, mean)
nameMatchVar14 = setNames(as.character(averageTv$tv), averageTv$componentNumber)
averageTvDF = data.frame(lapply(compDF, function(i) nameMatchVar14[i]))
averageTvOfComponent = averageTvDF$components.g..membership
averageTheater = aggregate(theater ~ componentNumber, data=dat2, mean)
nameMatchVar15 = setNames(as.character(averageTheater$theater), averageTheater$componentNumber)
averageTheaterDF = data.frame(lapply(compDF, function(i) nameMatchVar15[i]))
averageTheaterOfComponent = averageTheaterDF$components.g..membership
averageMovies = aggregate(movies ~ componentNumber, data=dat2, mean)
nameMatchVar16 = setNames(as.character(averageMovies$movies), averageMovies$componentNumber)
averageMoviesDF = data.frame(lapply(compDF, function(i) nameMatchVar16[i]))
averageMoviesOfComponent = averageMoviesDF$components.g..membership
averageConcerts = aggregate(concerts ~ componentNumber, data=dat2, mean)
nameMatchVar17 = setNames(as.character(averageConcerts$concerts), averageConcerts$componentNumber)
averageConcertsDF = data.frame(lapply(compDF, function(i) nameMatchVar17[i]))
averageConcertsOfComponent = averageConcertsDF$components.g..membership
averageMusic = aggregate(music ~ componentNumber, data=dat2, mean)
nameMatchVar18 = setNames(as.character(averageMusic$music), averageMusic$componentNumber)
averageMusicDF = data.frame(lapply(compDF, function(i) nameMatchVar18[i]))
averageMusicOfComponent = averageMusicDF$components.g..membership
averageShopping = aggregate(shopping ~ componentNumber, data=dat2, mean)
nameMatchVar19 = setNames(as.character(averageShopping$shopping), averageShopping$componentNumber)
averageShoppingDF = data.frame(lapply(compDF, function(i) nameMatchVar19[i]))
averageShoppingOfComponent = averageShoppingDF$components.g..membership
averageYoga = aggregate(yoga ~ componentNumber, data=dat2, mean)
nameMatchVar20 = setNames(as.character(averageYoga$yoga), averageYoga$componentNumber)
averageYogaDF = data.frame(lapply(compDF, function(i) nameMatchVar20[i]))
averageYogaOfComponent = averageYogaDF$components.g..membership

# change to numeric
dat2$componentTotalNodes = as.numeric(paste(componentTotalNodes))
dat2$averageAgeOfComponent = round(as.numeric(paste(averageAgeOfComponent)))
dat2$percentOfMalesInComponent = round(as.numeric(paste(percentOfMalesInComponent)),3)
dat2$averageImpraceOfComponent = round(as.numeric(paste(averageImpraceOfComponent)),3)
dat2$averageSportsOfComponent = round(as.numeric(paste(averageSportsOfComponent)),3)
dat2$averageTvsportsOfComponent = round(as.numeric(paste(averageTvsportsOfComponent)),3)
dat2$averageExerciseOfComponent = round(as.numeric(paste(averageExerciseOfComponent)),3)
dat2$averageDiningOfComponent = round(as.numeric(paste(averageDiningOfComponent)),3)
dat2$averageMuseumsOfComponent = round(as.numeric(paste(averageMuseumsOfComponent)),3)
dat2$averageHikingOfComponent = round(as.numeric(paste(averageHikingOfComponent)),3)
dat2$averageGamingOfComponent = round(as.numeric(paste(averageGamingOfComponent)),3)
dat2$averageClubbingOfComponent = round(as.numeric(paste(averageClubbingOfComponent)),3)
dat2$averageReadingOfComponent = round(as.numeric(paste(averageReadingOfComponent)),3)
dat2$averageTvOfComponent = round(as.numeric(paste(averageTvOfComponent)),3)
dat2$averageTheaterOfComponent = round(as.numeric(paste(averageTheaterOfComponent)),3)
dat2$averageMoviesOfComponent = round(as.numeric(paste(averageMoviesOfComponent)),3)
dat2$averageConcertsOfComponent = round(as.numeric(paste(averageConcertsOfComponent)),3)
dat2$averageMusicOfComponent = round(as.numeric(paste(averageMusicOfComponent)),3)
dat2$averageShoppingOfComponent = round(as.numeric(paste(averageShoppingOfComponent)),3)
dat2$averageYogaOfComponent = round(as.numeric(paste(averageYogaOfComponent)),3)
```

A seed was set so that this report would stay consistent. Testing was done without setting a seed.

```{r}
set.seed(231)
```

The data was split into training and test data with a 4:1 ratio.

```{r}
# get test and training sets
splits = split_data(dat2, frac=c(4,1))
tr_data = splits[[1]]
te_data = splits[[2]]
```

### Predictions using Logistic Regression

This particular model was made using age, gender, imprace, sports, exercise, gaming, clubbing, and tv as predictors to determine if they had a successful match. Imprace rated on a scale from 1-10 on how important it is that they date someone of the same race. The activites used as features after imprace indicate how interested an individual is in those activities

```{r}
fit = glm(hasMatch ~ age + gender + imprace + sports + exercise + gaming + clubbing + tv, data=tr_data, family=binomial)
```

Logistic Regression was used here with a threshold of >0.7. A confusion matrix was created to check the accuracy.

```{r}
# compute confusion matrix
y = predict(fit, newdata=te_data, type="response")
predicts = as.numeric(y > 0.7)
actuals = te_data$hasMatch
conf_mtx = table(predicts, actuals)
conf_mtx
mean(predicts == actuals)
```

This model was made with the same predictors as the first model except it also uses the total number of nodes in the component they belong in.

```{r}
roc_obj = roc(actuals, y)
auc(roc_obj)
plot(roc_obj, main="ROC", xlim=c(1,0), ylim=c(0,1), col="red4")
```

```{r}
fit2 = glm(hasMatch ~ averageAgeOfComponent + percentOfMalesInComponent + averageImpraceOfComponent + averageSportsOfComponent + averageExerciseOfComponent + averageGamingOfComponent + averageClubbingOfComponent + averageTvOfComponent + componentTotalNodes, data=tr_data, family=binomial)
```

Logistic Regression was used here as well with the same threshold.

```{r}
# compute confusion matrix
y2 = predict(fit2, newdata=te_data, type="response")
predicts = as.numeric(y2 > 0.7)
actuals = te_data$hasMatch
conf_mtx = table(predicts, actuals)
conf_mtx
mean(predicts == actuals)
```

```{r}
roc_obj2 = roc(actuals, y2)
auc(roc_obj2)
plot(roc_obj2, main="ROC", xlim=c(1,0), ylim=c(0,1), col="red4")
```

### Predictions using Classification Trees

We tried to make predictions using the same predictors that were used in the Logistic Regression predictions.

```{r}
fit = rpart(hasMatch ~ age + gender + imprace + sports + exercise + gaming + clubbing + tv, dat = tr_data, method = "class", control=rpart.control(minsplit=5, minbucket=3, cp=0.01))
prp(fit, extra=106, varlen= 0,faclen=0, main = "", box.col = c("paleturquoise", "pink")[fit$frame$yval])
```

This is the accuracy for the Classification Tree using node attributes.

```{r}
predicted = predict(fit, te_data, type = "class")
actual = te_data$hasMatch
table(actual, predicted)
mean(actual == predicted)
```

The AUC of this model is terrible and under .50.

```{r}
roc_obj = roc(actual, as.numeric(predicted) - 1)
auc(roc_obj)
plot(roc_obj, main="ROC", xlim=c(1,0), ylim=c(0,1), col="red4")
```

This is the second Classification Tree that uses graph component aggregates.

```{r}
fit2 = rpart(hasMatch ~ averageAgeOfComponent + percentOfMalesInComponent + averageImpraceOfComponent + averageSportsOfComponent + averageExerciseOfComponent + averageGamingOfComponent + averageClubbingOfComponent + averageTvOfComponent + componentTotalNodes, dat = tr_data, method = "class", control=rpart.control(minsplit=1, minbucket=1, cp=0.01))
prp(fit2, extra=106, varlen= 0,faclen=0, main = "", box.col = c("paleturquoise", "pink")[fit$frame$yval])
```

The accuracy for this model using a confusion matrix is the same as the above model.

```{r}
predicted = predict(fit2, te_data, type = "class")
actual = te_data$hasMatch
table(actual, predicted)
mean(actual == predicted)
```

The AUC for this model is also fairly poor as it is under .60.

```{r}
roc_obj2 = roc(actual, as.numeric(predicted) - 1)
auc(roc_obj2)
plot(roc_obj2, main="ROC", xlim=c(1,0), ylim=c(0,1), col="red4")
```

### Conclusion

From what could be seen, the AUC for the prediction that uses graph component aggregates appears to be commonly higher in the case of Logistic Regression. This has been tested with a couple of different seeds and there were only a couple of occurrances where the AUC for using node information instead of graph aggregates had a higher AUC. There are also many more features that can be used when creating these models. Classification Trees were not very effective when making predictions however that can be because of this particular data set. This shows that using the AUC has a higher potential of being more accurate and effective in making predictions when using Logistic Regression.