---
title: "Prediction Algorithm using Network Statistics"
author: "David Jia, David Keller, Alvaro Gonzalez"
date: "May 2, 2018"
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(prompt=TRUE, comment="", echo=TRUE)
```

The igraph library is necessary.

```{r collapse=TRUE, warning=FALSE}
library(igraph)
```

The data set used was a Speed Dating one.

```{r}
dat = read.csv("SpeedDatingData.csv")
```

```{r, echo=FALSE}
# add negative one imputation
impute_negative_one = function(x) {
  stopifnot(length(x) > 0 && sum(!is.na(x)) > 0)
  x[is.na(x)] = -1
  return(x)
}

# add mean imputation
impute_mean = function(x) {
  stopifnot(length(x) > 0 && sum(!is.na(x)) > 0)
  x[is.na(x)] = median(x, na.rm=TRUE)
  return(x)
}

split_data = function(dat, frac=c(0.75, 0.25)) {
  # at least one set must be specified
  k = length(frac)
  stopifnot(k > 0)
  
  n = nrow(dat)
  frac = frac/(sum(frac))
  starts = c(1, round(cumsum(frac) * n)[-k])
  ends = c(starts[-1]-1,n)
  samp = sample(1:n)
  data_sets = list()
  for (i in 1:k) {
    data_sets[[i]] = dat[samp[starts[i]:ends[i]],]
  }
  return(data_sets)
} 

layout.by.attr <- function(graph, wc, cluster.strength=1,layout=layout.auto) {  
  g <- graph.edgelist(get.edgelist(graph)) # create a lightweight copy of graph w/o the attributes.
  E(g)$weight <- 1
  
  attr <- cbind(id=1:vcount(g), val=wc)
  g <- g + vertices(unique(attr[,2])) + igraph::edges(unlist(t(attr)), weight=cluster.strength)
  
  l <- layout(g, weights=E(g)$weight)[1:vcount(graph),]
  return(l)
}
```

### Creating Second Data Set

Filling in some of the missing data from the original data set and create a second data set that has the attribute of each individual node. The orgiinal data set had attributes from the pairings of two nodes.

```{r}
dat$age = impute_mean(dat$age)
dat$age_o = impute_mean(dat$age_o)
dat$race = impute_negative_one(dat$race)

dat2 = data.frame(iid=unique(dat$iid))
dat2$age = aggregate(age ~ iid, data = dat, mean)$age
dat2$gender = aggregate(gender ~ iid, data = dat, mean)$gender
dat2$race = aggregate(race ~ iid, data = dat, mean)$race
dat2$hasMatch = aggregate(match ~ iid, data = dat, max)$match
dat2$metWithSameRace = aggregate(samerace ~ iid, data = dat, max)$samerace
```

```{r, echo=FALSE}
iidOfSameRace = dat[dat$samerace == 1,]$iid
pidOfSameRace = dat[dat$samerace == 1,]$pid

edges = c()

i = 0
while (i < length(iidOfSameRace)) {
  iid = iidOfSameRace[i+1]
  pid = pidOfSameRace[i+1]
  if (iid > 117) {
    iid = iid - 1
  }
  if (is.na(pid)) {
    pid = iid
  }
  if (pid > 117 & !is.na(pid)) {
    pid = pid - 1
  }
  edges = c(edges, iid)
  edges = c(edges, pid)
  i = i + 1
}
```

### Graph

The graph created here shows the connections between nodes where they met with someone of the same race.

```{r, echo=FALSE}
g = make_empty_graph(n = length(unique(dat$iid)), directed=FALSE) %>% add_edges(edges)
g = simplify(g, remove.multiple=TRUE)
V(g)$label = unique(dat$iid)
plot(g, vertex.size=2, edge.width=1, edge.color="red", vertex.label.cex=0.5, vertex.label.dist=0.4, vertex.label.degree=-pi/2, layout=layout.by.attr(g, wc=1))
```

The number of nodes in each component was recorded and assigned to the respective nodes in the data set.

```{r}
compDF = data.frame(components(g)$membership)
compTableDF = data.frame(table(components(g)$membership))
nameMatchVar = setNames(as.character(compTableDF$Freq), compTableDF$Var1)
totalCompDF = data.frame(lapply(compDF, function(i) nameMatchVar[i]))
componentTotalNodes = totalCompDF$components.g..membership

# change to numeric
dat2$componentTotalNodes = as.numeric(paste(componentTotalNodes))
```

A seed was set so that this report would stay consistent. Testing was done without setting a seed.

```{r}
set.seed(321)
```

The data was split into training and test data with a 4:1 ratio.

```{r}
# get test and training sets
splits = split_data(dat2, frac=c(4,1))
tr_data = splits[[1]]
te_data = splits[[2]]
```

### Predictions

This particular model was made using age, gender, and if they met with someone from the same race as predictors to determine if they had a successful match.

```{r}
fit = glm(hasMatch ~ age + gender + metWithSameRace, data=tr_data, family=binomial)
```

Logistic Regression was used here with a threshold of >0.7. A confusion matrix was created to check the accuracy.

```{r}
# compute confusion matrix
y = predict(fit, newdata=te_data, type="response")
predicts = as.numeric(y > 0.7)
actuals = te_data$hasMatch
conf_mtx = table(predicts, actuals)
conf_mtx
mean(predicts == actuals)
```

This model was made with the same predictors as the first model except it also uses the total number of nodes in the component they belong in.

```{r}
fit2 = glm(hasMatch ~ age + gender + metWithSameRace + componentTotalNodes, data=tr_data, family=binomial)
```

Logistic Regression was used here as well with the same threshold.

```{r}
# compute confusion matrix
y = predict(fit2, newdata=te_data, type="response")
predicts = as.numeric(y > 0.7)
actuals = te_data$hasMatch
conf_mtx = table(predicts, actuals)
conf_mtx
mean(predicts == actuals)
```

### Conclusion